Investigation of Limited Detection Schemes for Light Scattering of Optically Trapped Asymmetric Particles

### 2023-08-22 Manuscript/Manuscript_Colloids.tex
linux pc push

\begin{figure} [b]
\centering
\begin{subfigure}{0.45\textwidth}
	\subcaption{}
	\includegraphics[width=\textwidth, height=0.25\textheight]{./Images/fig1a.png}
\end{subfigure}
\begin{subfigure}{0.45\textwidth}
	\subcaption{}
	\includegraphics[width=\textwidth, height=0.25\textheight]{./Images/fig1b.png}
\end{subfigure}
\caption{\label{fig:setup}
  %
  Proposed experimental set up for scattering measurements from an object in an optical trap. The probe beam for scattering measurements is incident perpendicular to the trapping laser propagation direction. a) Side view. b) Top view. Note that three of the detector fibres are co-planar with the incident probe beam, while the fourth detector is placed out of the plane (see Sec~\ref{sec:detectors}).
%
}
\end{figure}

@@

\begin{figure}[h]
	\centering
	\begin{subfigure}{0.49\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/fig2a.png}
	\end{subfigure}
	\begin{subfigure}{0.49\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/fig2b.png}
	\end{subfigure}
	\caption{(a)Example dimer in orientation $\bar{\bf s}$, (b) 30 Reference orientations represented by vectors pointing from [0,0,0] to each point}
	\label{fig:dimer}
\end{figure}

@@

\begin{figure}[h]
	\centering
	\begin{subfigure}{0.4\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/fig3a.png}
	\end{subfigure}
	\begin{subfigure}{0.4\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/fig3b.png}
	\end{subfigure}
	\caption{(a) Distribution of orientation vectors and (b) their respective scattering signals. Points are coloured according to their distance from the centre of each cluster (red points centred around [$0.00, 0.00, 1.00$], blue points centred at [$0.71, 0.00, 0.71$])}
	\label{fig:mixing}
\end{figure}

@@

\begin{figure}[h]
	\centering
	\begin{subfigure}{0.49\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/fig4a.png}
	\end{subfigure}
	\begin{subfigure}{0.4\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/fig4b.png}
	\end{subfigure}
	\caption{Expected scattered signals from reference orientations -see fig~\ref{fig:dimer} - when: (a) all three detectors are in the X-Y plane, (b) when 1 detector is raised out of the X-Y plane.}
	\label{fig:detectors}
\end{figure}

@@

\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth]{./Images/fig5.png}
	\caption{Model's estimation of dimer orientation over the simulation time, assuming uniform prior $p(\hat{\bf n}_\alpha)$, broken up into x, y, and z components for clarity. Blue line denotes the best result we can achieve (the reference orientation $\hat{\bf n}_{best}$ that is closest to the actual orientation), orange line denotes the result provided by eq~\ref{eq:bayes}: where the orange line is not visible, the model's prediction agrees with $\hat{\bf n}_{best}$. Dotted black line is the instantaneous orientation $\hat{\bf s}$.}
	\label{fig:uniform}
\end{figure}

@@

\begin{figure}[h]
\centering
\includegraphics[width=0.5\textwidth]{./Images/fig8a.png}
\caption{\label{fig:biased}
%
Estimation of dimer orientation with $p(\hat{\bf n}_\alpha)$ defined
by Eq~\eqref{eq:boltz}.  Blue line denotes the best result we can
achieve, orange line denotes the result provided by eq~\ref{eq:bayes}.
Dotted black line is the instantaneous orientation $\hat{\bf s}$ (see Section~\ref{sec:Bayes}).
  %
}
\end{figure}

@@

\begin{figure}[h]
	\centering
	\begin{subfigure}{0.45\textwidth}
		\subcaption{}
		\includegraphics[width =\textwidth]{./Images/fig7a.png}
	\end{subfigure}
	\begin{subfigure}{0.45\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/fig7b.png}
	\end{subfigure}
	\caption{Simulation results of: (a) the dimer's orientation vector with time, (b) the dimer's [x,y,z] position with time.}
	\label{fig:motion}
\end{figure}

@@

\begin{figure}[h]
	\centering
	\begin{subfigure}{0.33\textwidth}
			\subcaption{}
			\includegraphics[width =\textwidth]{./Images/fig8a.png}
	\end{subfigure}
	\begin{subfigure}{0.31\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/fig8b.png}
	\end{subfigure}
	\begin{subfigure}{0.31\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/fig8c.png}
	\end{subfigure}
	\caption{Model estimates of orientation when neural net has been trained on dimer of size ratio: (a) 1:2 $[F(K_l)=9.456]$, (b) 1:2.05 $[F(K_l)=1.324]$, (c) 1:1.95 $[F(K_l)=1.325]$ ($n_{refs} = 30$)}
	\label{fig:size}
\end{figure}
As can be seen from Fig~\ref{fig:size}  even the slightest change in size ratio makes a very significant difference to the performance of our model. This amounts to just over $100 nm$ in the dimer's overall size, yet results in our model being correct from over 90 \% of the time to now as low as 30 \%. This highlights the importance of correctly sizing trapped entities before performing any in depth analysis of the scattering pattern, as even the slightest deviation can have a serious impact. We addressed this by increasing the number of available reference points from 30 to 126 (following the same procedure as given by \cite{Rey2006} to evenly space out the coordinates) and increasing the weighting factor in Eq~\ref{eq:boltz}. While this didn't have a significant improvement on the overall accuracy of the model, in the worst case having a slight increase from 30.5 \% to 40.3 \%, it did help to significantly reduce the magnitude between our model's estimations and the dimer's motion as seen below in Fig~\ref{fig:refs}.
\begin{figure}[h]
	\centering
	\begin{subfigure}{0.31\textwidth}
		\subcaption{}
		\includegraphics[width =\textwidth]{./Images/fig9a.png}
	\end{subfigure}
	\begin{subfigure}{0.31\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/fig9b.png}
	\end{subfigure}
	\begin{subfigure}{0.31\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/fig9c.png}
	\end{subfigure}
	\caption{Model estimates of orientation when neural net has been trained on dimer of size ratio: (a) 1:2 $[F(K_l)=11.756]$, (b) 1:2.05 $[F(K_l)=1.233]$, (c) 1:1.95 $[F(K_l)=2.128]$, ($n_{refs} = 126$)}
	\label{fig:refs}
\end{figure}

@@

\begin{figure}[h]
	\centering
	\begin{subfigure}{0.32\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/fig10a.png}
	\end{subfigure}
	\begin{subfigure}{0.32\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/fig10b.png}
	\end{subfigure}
	\begin{subfigure}{0.32\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/fig10c.png}
	\end{subfigure}
	\caption{Model prediction for signal error of (a) $1\%$ ~  $[F(K_l)=7.246]$, (b) $15\%$ $[F(K_l)=0.511]$, and (c) $25\%$ ~  $[F(K_l)=0.536]$.}
	\label{fig:epsilon}
\end{figure}

@@

To reduce the effects of the signal noise we took the time average of the expected signal over 0.02s and then had our neural network estimate the orientation based on the average signal.
\begin{figure}[h]
	\centering
	\begin{subfigure}{0.32\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/fig11a.png}
	\end{subfigure}
	\begin{subfigure}{0.32\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/fig11b.png}
	\end{subfigure}
	\begin{subfigure}{0.32\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/fig11c.png}
	\end{subfigure}
	\caption{Model prediction for signal error of (a) $1\%$ $[F(K_l)=4.823]$,  (b) $15\%$ $[F(K_l)=1.494]$, and (c) $25\%$ $[F(K_l)=0.882]$, time averaged over $20\,{\rm ms}$}
	\label{fig:time average}
\end{figure}
This resulted in a reduction in the overall signal noise and provided a higher degree of accuracy for our model. There appears to be no clear correlation between the length over which we time average and the performance of our model. Time averaging over every 0.05s resulted in a drastically worse performance; this is due to the fact that over longer time periods there is greater uncertainty regarding how the dimer's orientation has changed, thus tracking the instantaneous orientation becomes harder for the neural network. Fortunately, time averaging even over $20\,{\rm ms}$ seems to provide a satisfactory estimation of the dimer's angular dynamics within the optical trap.

>>>>

\begin{center}
		{\bf{Fig 1a, 1b}}
\end{center}
\begin{figure} [b]
\centering
\begin{subfigure}{0.45\textwidth}
	\subcaption{}
	\includegraphics[width=\textwidth, height=0.25\textheight]{./Images/fig1a.png}
\end{subfigure}
\begin{subfigure}{0.45\textwidth}
	\subcaption{}
	\includegraphics[width=\textwidth, height=0.25\textheight]{./Images/fig1b.png}
\end{subfigure}
\caption{\label{fig:setup}
  %
  Proposed experimental set up for scattering measurements from an object in an optical trap. The probe beam for scattering measurements is incident perpendicular to the trapping laser propagation direction. a) Side view. b) Top view. Note that three of the detector fibres are co-planar with the incident probe beam, while the fourth detector is placed out of the plane (see Sec~\ref{sec:detectors}).
%
}
\end{figure}

@@

\begin{center}
	\bf{Fig 2a, 2b}
\end{center}
\begin{figure}[h]
	\centering
	\begin{subfigure}{0.49\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/fig2a.png}
	\end{subfigure}
	\begin{subfigure}{0.49\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/fig2b.png}
	\end{subfigure}
	\caption{(a)Example dimer in orientation $\bar{\bf s}$, (b) 30 Reference orientations represented by vectors pointing from [0,0,0] to each point}
	\label{fig:dimer}
\end{figure}

@@

\begin{center}
	\bf{Fig 3a, 3b}
\end{center}
\begin{figure}[h]
	\centering
	\begin{subfigure}{0.4\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/fig3a.png}
	\end{subfigure}
	\begin{subfigure}{0.4\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/fig3b.png}
	\end{subfigure}
	\caption{(a) Distribution of orientation vectors and (b) their respective scattering signals. Points are coloured according to their distance from the centre of each cluster (red points centred around [$0.00, 0.00, 1.00$], blue points centred at [$0.71, 0.00, 0.71$])}
	\label{fig:mixing}
\end{figure}

@@

\begin{center}
	\bf{Fig 4a, 4b}
\end{center}
\begin{figure}[h]
	\centering
	\begin{subfigure}{0.49\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/fig4a.png}
	\end{subfigure}
	\begin{subfigure}{0.4\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/fig4b.png}
	\end{subfigure}
	\caption{Expected scattered signals from reference orientations -see fig~\ref{fig:dimer} - when: (a) all three detectors are in the X-Y plane, (b) when 1 detector is raised out of the X-Y plane.}
	\label{fig:detectors}
\end{figure}

@@

\begin{center}
	\bf{Fig 5}
\end{center}
\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth]{./Images/fig5.png}
	\caption{Model's estimation of dimer orientation over the simulation time, assuming uniform prior $p(\hat{\bf n}_\alpha)$, broken up into x, y, and z components for clarity. Blue line denotes the best result we can achieve (the reference orientation $\hat{\bf n}_{best}$ that is closest to the actual orientation), orange line denotes the result provided by eq~\ref{eq:bayes}: where the orange line is not visible, the model's prediction agrees with $\hat{\bf n}_{best}$. Dotted black line is the instantaneous orientation $\hat{\bf s}$.}
	\label{fig:uniform}
\end{figure}

@@

\begin{center}
	\bf{Fig 6}
\end{center}
\begin{figure}[h]
\centering
\includegraphics[width=0.5\textwidth]{./Images/fig8a.png}
\caption{\label{fig:biased}
%
Estimation of dimer orientation with $p(\hat{\bf n}_\alpha)$ defined
by Eq~\eqref{eq:boltz}.  Blue line denotes the best result we can
achieve, orange line denotes the result provided by eq~\ref{eq:bayes}.
Dotted black line is the instantaneous orientation $\hat{\bf s}$ (see Section~\ref{sec:Bayes}).
  %
}
\end{figure}

@@

\begin{center}
	\bf{Fig 7a, 7b}
\end{center}
\begin{figure}[h]
	\centering
	\begin{subfigure}{0.45\textwidth}
		\subcaption{}
		\includegraphics[width =\textwidth]{./Images/fig7a.png}
	\end{subfigure}
	\begin{subfigure}{0.45\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/fig7b.png}
	\end{subfigure}
	\caption{Simulation results of: (a) the dimer's orientation vector with time, (b) the dimer's [x,y,z] position with time.}
	\label{fig:motion}
\end{figure}

@@

\begin{center}
	\bf{Fig 8a, 8b, 8c}
\end{center}
\begin{figure}[h]
	\centering
	\begin{subfigure}{0.33\textwidth}
			\subcaption{}
			\includegraphics[width =\textwidth]{./Images/fig8a.png}
	\end{subfigure}
	\begin{subfigure}{0.31\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/fig8b.png}
	\end{subfigure}
	\begin{subfigure}{0.31\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/fig8c.png}
	\end{subfigure}
	\caption{Model estimates of orientation when neural net has been trained on dimer of size ratio: (a) 1:2 $[F(K_l)=9.456]$, (b) 1:2.05 $[F(K_l)=1.324]$, (c) 1:1.95 $[F(K_l)=1.325]$ ($n_{refs} = 30$)}
	\label{fig:size}
\end{figure}
As can be seen from Fig~\ref{fig:size}  even the slightest change in size ratio makes a very significant difference to the performance of our model. This amounts to just over $100 \ nm$ in the dimer's overall size, yet results in our model being correct from over 90 \% of the time to now as low as 30 \%. This highlights the importance of correctly sizing trapped entities before performing any in depth analysis of the scattering pattern, as even the slightest deviation can have a serious impact. We addressed this by increasing the number of available reference orientations from 30 to 126 (following the same procedure as given by \cite{Rey2006} to evenly space out the coordinates) and increasing the weighting factor in Eq~\ref{eq:boltz}. While this didn't have a significant improvement on the overall accuracy of the model, in the worst case having a slight increase from 30.5 \% to 40.3 \%, it did help to significantly reduce the magnitude between our model's estimations and the dimer's motion as seen below in Fig~\ref{fig:refs}.

\begin{center}
	\bf{Fig 9a, 9b, 9c}
\end{center}
\begin{figure}[h]
	\centering
	\begin{subfigure}{0.31\textwidth}
		\subcaption{}
		\includegraphics[width =\textwidth]{./Images/fig9a.png}
	\end{subfigure}
	\begin{subfigure}{0.31\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/fig9b.png}
	\end{subfigure}
	\begin{subfigure}{0.31\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/fig9c.png}
	\end{subfigure}
	\caption{Model estimates of orientation when neural net has been trained on dimer of size ratio: (a) 1:2 $[F(K_l)=11.756]$, (b) 1:2.05 $[F(K_l)=1.233]$, (c) 1:1.95 $[F(K_l)=2.128]$, ($n_{refs} = 126$)}
	\label{fig:refs}
\end{figure}

@@

\begin{center}
	\bf{Fig 10a, 10b, 10c}
\end{center}
\begin{figure}[h]
	\centering
	\begin{subfigure}{0.32\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/fig10a.png}
	\end{subfigure}
	\begin{subfigure}{0.32\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/fig10b.png}
	\end{subfigure}
	\begin{subfigure}{0.32\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/fig10c.png}
	\end{subfigure}
	\caption{Model prediction for signal error of (a) $1\%$ ~  $[F(K_l)=7.246]$, (b) $15\%$ $[F(K_l)=0.511]$, and (c) $25\%$ ~  $[F(K_l)=0.536]$.}
	\label{fig:epsilon}
\end{figure}

@@

To reduce the effects of the signal noise we took the time average of the expected signal over $0.001 s$ and then had our neural network estimate the orientation based on the average signal.

\begin{center}
	\bf{Fig 11a, 11b, 11c}
\end{center}
\begin{figure}[h]
	\centering
	\begin{subfigure}{0.32\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/fig11a.png}
	\end{subfigure}
	\begin{subfigure}{0.32\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/fig11b.png}
	\end{subfigure}
	\begin{subfigure}{0.32\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/fig11c.png}
	\end{subfigure}
	\caption{Model prediction for signal error of (a) $1\%$ $[F(K_l)=4.823]$,  (b) $15\%$ $[F(K_l)=1.494]$, and (c) $25\%$ $[F(K_l)=0.882]$, time averaged over $1\,{\rm ms}$}
	\label{fig:time average}
\end{figure}
This resulted in a reduction in the overall signal noise and provided a higher degree of accuracy for our model. There appears to be no clear correlation between the length over which we time average and the performance of our model. Time averaging over every 0.05s resulted in a drastically worse performance; this is due to the fact that over longer time periods there is greater uncertainty regarding how the dimer's orientation has changed, thus tracking the instantaneous orientation becomes harder for the neural network. Fortunately, time averaging even over $1\,{\rm ms}$ seems to provide a satisfactory estimation of the dimer's angular dynamics within the optical trap.
